---
title: "Evaluating Machine Learning Algorithms"
output: html_notebook
---

> This is a tutorial provided by machinelearningmastery.com. Walking through the tutorial is intended to provide me with a better understanding of the machine learning process.

> The purpose of this tutorial is determining which algorithm provides the most accurate model for the data set.The method describe is the use of trial and error.

> There are 3 sections to the tutorial.
+ Defining a test harness
+ Building multiple predictive models from the data
+ Comparing models and selecting a short list

**Test Harness**

> The test harness is comprised of three key elements.
1. The dataset we will use to train models.
2. The test options used to evaluate a model (resampling method)
3. The metric we are interested in measuring and comparing.

> Rule of thumb when spot checking is that each algorithm should train within 1-to-2 minutes. Less than 10,000 rows is a good dataset size.

> Start by loading up the libraries and the dataset.

```{r}
# load libraries
library(mlbench)
library(caret)

# assign filepath
filepath <- paste(getwd(), "/data/pima-indians-diabetes.csv", sep="")
# load data
dataset <- read.csv(filepath, header = FALSE)
```

> At this point the tutorial jumps right into test options and skips a couple steps. For the sake of developing my understanding I will be inserting a couple steps. The first step is to partition the data into a training data set and a validation data set. The training data set is used to train the model and the validation data set will be used to check whether the chosen model works.

```{r}
set.seed(123)

validation_index <- createDataPartition(y = dataset$V9, p = .8, list = FALSE)

# Assign partions
training <- dataset[validation_index, ]
validation <- dataset[-validation_index, ]

nrow(training)

nrow(validation)
```
