---
title: "Caret Package Walkthrough"
output: html_notebook
---

##Loading the Dataset##

> First thing is first, we need to load the data. This walkthrough introduces the **mlbench** package. Mlbench is a collection of artificial and real-world machine learning benchmark problems, including several data sets from the UCI repository. Which is great. This just makes life easier when practicing machine learning.

> This walkthrough uses the **Sonar** data set. It consists of 208 data points collected on 60 predictors. The goal is to predict the two classes **M** for metal cylinder or **R** for cylinder shaped rock.

**Split the Data into Two Groups**

```{r}
library(caret)
library(mlbench)
data(Sonar)

set.seed(107)
inTrain <- createDataPartition(
  y = Sonar$Class,
  ## the outcome data are needed
  p = .75,
  ## the percentage of data in the training set
  list = FALSE
)
## The format of the results

## The output is a set of integers fro the rows of Sonar
## that belong in the training set.
str(inTrain)
#> int [1:157, 1] 1 2 3 4 5 7
#> - attr(*, "dimnames")=list of 2
#> ..$ : NULL
#> ..$ : chr "Resample1"
```

> **createDataPartition** does a stratified random split of the data.

```{r}
training <- Sonar[ inTrain, ]
testing  <- Sonar[-inTrain, ]

nrow(training)
#> [1] 157
nrow(testing)
#> [1] 51
```

> Next a partial least squares discriminant analysis (PLSDA) model will be tudned over the number of PLS components that should be retained.

```{r}
plsFit <- train( Class ~., data = training, method = "pls",
                 ## Center and scale the predictors for the training
                 ## set and all future samples.
                 preProc = c("center", "scale")
                 )
```


> At this point the model can be customized. The walkthrough recommends expanding the PLS models, changing the type of resample from bootstrap to three repeats of 10-fold cross-validation and changing the methods for measuring performance. Normally the root mean square error and R^2 are computed, but it will need to change to estimate the area under the ROC curve. 

